---
title: "Analyses Discussed in Footnote 6"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: default
---

## (1) Description of Data File

```{r include=FALSE}
options(width=100)

library(RCurl)
library(data.table)
library(pequod)
library(psych)
library(rtf)

## DATA PREPARATION ##

# import data
csv <- getURL("https://raw.githubusercontent.com/abrowman/climate-pspb2016/master/Footnote%206/Climate_Footnote6.csv")
mydata <- read.csv(textConnection(csv))
```

The posted data for this study consists of 1 compiled, deidentified data file. `r nrow(mydata)` participants completed the study, but `r nrow(subset(mydata, mydata$Class >= 5))` were excluded because they were not undergraduate students.

```{r include=FALSE}
mydata <- subset(mydata, Class < 5)
attach(mydata)

# score measures
cond <- factor(ifelse(!is.na(Ntrl1), "chilly", ifelse(!is.na(Ntrl3), "warm" ,NA)), levels=c("warm", "chilly"))
cond.num <- ifelse(cond=="chilly", -1, ifelse(cond=="warm", 1, NA))
options(contrasts=c("contr.sum", "contr.poly"))
ssid <- (PSF1r+PSF2r+(8-PSF3)+PSF4r+PSF5r+(8-PSF6))/6
beluncert <- (Belonging2+Belonging3)/2
belong <- Belonging1

# center continuous predictors
income_c <- scale(Income,scale=F,center=T)
ladder_c <- scale(USldr,scale=F,center=T)
```

## (2) Descriptive Statistics

**M~age~ (*SD*):** `r round(mean(Age, na.rm=T), digits=1)` (`r round(sd(Age, na.rm=T), digits=2)`)

**Percentage by gender:**
```{r echo=FALSE}
gender <- factor(ifelse(Gender==1, "male", ifelse(Gender==2, "female", NA)))
round(sort((table(gender, useNA="always")/nrow(mydata)*100),decreasing=F), digits=1)
```

**Percentage by class:**
```{r echo=FALSE}
class <- ifelse(Class==1, "Freshman", ifelse(Class==2, "Sophomore", ifelse(Class==3, "Junior", ifelse(Class==4, "Senior", NA))))
round(sort((table(class, useNA="always")/nrow(mydata))*100,decreasing=T), digits=1)
```

**Percentage by income:**
```{r echo=FALSE}
income.label <- ifelse(Income==1, "(1) $25,000 or less", ifelse(Income==2, "(2) $25,001-$40,000", ifelse(Income==3, "(3) $40,001-$70,000", ifelse(Income==4, "(4) $70,001-$90,000", ifelse(Income==5, "(5) $90,001-$120,000", ifelse(Income==6, "(6) $120,001-$150,000", ifelse(Income==7, "(7) $150,001-$200,000", ifelse(Income==8, "(8) $200,001-$300,000", ifelse(Income==9, "(9) $300,001 or more", NA)))))))))
round((table(income.label, useNA="always")/nrow(mydata))*100, digits=1)
```

***M*s and *SD*s for sensitivity to SES-based identity discrepancies (ssid), belonging uncertainty (beluncert), and belonging (belong):**  
```{r echo=FALSE}
describe(data.frame(Income, ssid, beluncert, belong))[,2:4]
```


**Alphas and within-scale correlations:**

  * **$\alpha$~SSID~** = `r round(alpha(data.frame(PSF1r, PSF2r, (8-PSF3), PSF4r, PSF5r, (8-PSF6)))$total$raw_alpha, digits=2)`  
  * ***r*~belonging-uncertainty~**(`r cor.test(Belonging2, Belonging3)$parameter`) = `r round(cor.test(Belonging2, Belonging3)$estimate, digits=2)`, *p* = `r signif(cor.test(Belonging2, Belonging3)$p.value, digits=2)`
  

## (3) Random Assignment Checks

**(a) Unequal variance test, *t*-test, and *M*s and *SD*s for: income**
```{r echo=FALSE}
var.test(Income ~ cond)
t.test(Income ~ cond, var.equal=TRUE)
describeBy(Income, cond)
```
\
**(b) Unequal variance test, *t*-test, and *M*s and *SD*s for: current GPA**
```{r echo=FALSE}
var.test(GPA ~ cond)
t.test(GPA ~ cond, var.equal=TRUE)
describeBy(GPA, cond)
```


## (4) Main Analyses

### SSID analyses

#### Omnibus test

**Estimates:**
```{r echo=FALSE}
summary(ssid.regress <- lm(ssid ~ income_c * cond.num))
```

**CIs:**
```{r echo=FALSE}
round(confint(ssid.regress, level = 0.95), digits=2)
```

***f*<sup>2</sup> value (effect size) for interaction term:** `r round((summary(ssid.regress)$r.squared-summary(lm(ssid ~ income_c + cond.num))$r.squared)/(1-summary(ssid.regress)$r.squared), digits=3)`

#### Simple slopes tests

**Simple slopes of condition among low- and high-SES students:**
```{r echo=FALSE}
contrasts(cond)
ssid.lmres <- lmres(ssid ~ income_c * cond.num, mydata)
summary(ssid.slopes <- simpleSlope(ssid.lmres, pred="cond.num", mod1="income_c"))
```

**CIs for simple slope among low-SES students:** `r round(ssid.slopes$simple_slope[1,1] + c(-1,1)*ssid.slopes$simple_slope[1,2]*qt(0.975, ssid.slopes$Df), digits=2)`  

**CIs for simple slope among high-SES students:** `r round(ssid.slopes$simple_slope[2,1] + c(-1,1)*ssid.slopes$simple_slope[2,2]*qt(0.975, ssid.slopes$Df), digits=2)`\
\


### Belonging uncertainty analyses

#### Omnibus test

**Estimates:**
```{r echo=FALSE}
summary(beluncert.regress <- lm(beluncert ~ income_c * cond.num))
```

**CIs:**
```{r echo=FALSE}
round(confint(beluncert.regress, level = 0.95), digits=2)
```

***f*<sup>2</sup> value (effect size) for interaction term:** `r round((summary(beluncert.regress)$r.squared-summary(lm(beluncert ~ income_c + cond.num))$r.squared)/(1-summary(beluncert.regress)$r.squared), digits=3)`


### Belonging analyses

#### Omnibus test

**Estimates:**
```{r echo=FALSE}
summary(belong.regress <- lm(belong ~ income_c * cond.num))
```

**CIs:**
```{r echo=FALSE}
round(confint(belong.regress, level = 0.95), digits=2)
```

***f*<sup>2</sup> value (effect size) for interaction term:** `r round((summary(belong.regress)$r.squared-summary(lm(belong ~ income_c + cond.num))$r.squared)/(1-summary(belong.regress)$r.squared), digits=3)`