---
title: "Experiment 3"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: default
---

## (1) Description of Data File

```{r include=FALSE}
options(width=100)

library(RCurl)
library(data.table)
library(pequod)
library(psych)
library(rtf)

## DATA PREPARATION ##

# import data
csv <- getURL("https://raw.githubusercontent.com/abrowman/climate-pspb2016/master/Experiment%203/Climate_Experiment3.csv")
mydata <- read.csv(textConnection(csv))
attach(mydata)
```

The posted data for this study consists of 1 compiled, deidentified data file (total *N* recruited = `r nrow(mydata)`). No data were excluded.

```{r include=FALSE}
# clean data
cgpa <- gsub("4.72", "3.72", cgpa)
cgpa <- gsub("3.6-3.7", "3.65", cgpa)
cgpa <- gsub("4.78", "3.78", cgpa)
cgpa <- as.numeric(gsub("Medical student", "", cgpa))
cond <- factor(ifelse(!is.na(wc1), "chilly", ifelse(!is.na(fd1), "warm", NA)), levels = c("warm", "chilly"))
cond.num <- ifelse(cond=="chilly", -1, ifelse(cond=="warm", 1, NA))
options(contrasts = c("contr.sum", "contr.poly"))

# compute scores
ladder <- 11-ladder.r
ssid <- (psc1r+psc2r+(8-psc3)+psc4r+psc5r+(8-psc6))/6
commit <- (sescommi1+sescommi2)/2
beluncert <- (beluncert1+beluncert3)/2
numrep <- sesdiverse_4
belong <- belong1

# center continuous predictors
income_c <- scale(income,scale=F,center=T)
ladder_c <- scale(ladder,scale=F,center=T)
```

## (2) Descriptive Statistics (Table 1)

***M*~age~ (*SD*):** `r round(mean(age, na.rm=T), digits=1)` (`r round(sd(age, na.rm=T), digits=2)`)

**Percentage by gender:**
```{r echo=FALSE}
Gender <- factor(ifelse(gender==1, "male", ifelse(gender==2, "female", NA)))
round((table(Gender, useNA="always")/nrow(mydata)*100), digits=1)
```

**Percentage by race:**  

*Note:* The percentage of Asians in the sample is slightly misreported in the article due to a typo. Correcting those values (as done here) does not change any conclusions or central values discussed in the paper.
```{r echo=FALSE}
Race <- ifelse(race==1, "Native-American", ifelse(race==2, "Asian", ifelse(race==3, "Black", ifelse(race==4, "Latino", ifelse(race==5, "Middle-Eastern", ifelse(race==6, "Islander", ifelse(race==7, "White", ifelse(race==8 | race=="7,8", "Multi-racial", NA)))))))) # sample by race: 1=native-american; 2=asian; 3=black; 4=latino; 5=middle eastern; 6=islander; 7=white; 8=multi-racial; blank=unreported
round(sort((table(Race, useNA="always")/nrow(mydata))*100,decreasing=T), digits=1)
```

**Percentage by class:**
```{r echo=FALSE}
Class <- ifelse(class==1, "Freshman", ifelse(class==2, "Sophomore", ifelse(class==3, "Junior", ifelse(class==4, "Senior", NA))))
round((table(Class, useNA="always")/nrow(mydata))*100, digits=1)
```

**Percentage by income:**
```{r echo=FALSE}
income.label <- ifelse(income==1, "(1) $25,000 or less", ifelse(income==2, "(2) $25,001-$40,000", ifelse(income==3, "(3) $40,001-$70,000", ifelse(income==4, "(4) $70,001-$90,000", ifelse(income==5, "(5) $90,001-$120,000", ifelse(income==6, "(6) $120,001-$150,000", ifelse(income==7, "(7) $150,001-$200,000", ifelse(income==8, "(8) $200,001-$300,000", ifelse(income==9, "(9) $300,001 or more", NA)))))))))
round((table(income.label, useNA="always")/nrow(mydata))*100, digits=1)
```

***M*s and *SD*s for income, numerical representation (numrep), perceived institutional commitment (commit), sensitivity to SES-based identity discrepancies (ssid), belonging (belong), belonging uncertainty (beluncert), and current GPA (cgpa):**  

*Note:* The *SD* for current GPA is slightly misreported in the published paper due to a typo. Correcting those values (as done here) does not change any conclusions or central values discussed in the paper.
```{r echo=FALSE}
describe(data.frame(income, numrep, commit, ssid, belong, beluncert, cgpa))[c(2,3,4)]
```

**Alphas and within-scale correlations:**

  * ***r*~perceived-institutional-commitment~**(`r cor.test(sescommi1, sescommi2)$parameter`) = `r round(cor.test(sescommi1, sescommi2)$estimate, digits=2)`, *p* = `r signif(cor.test(sescommi1, sescommi2)$p.value, digits=3)`
  * **$\alpha$~SSID~** = `r round(alpha(data.frame(psc1r, psc2r, (8-psc3), psc4r, psc5r, (8-psc6)))$total$raw_alpha, digits=2)`
  * ***r*~belonging-uncertainty~**(`r cor.test(beluncert1, beluncert3)$parameter`) = `r round(cor.test(beluncert1, beluncert3)$estimate, digits=2)`, *p* = `r signif(cor.test(beluncert1, beluncert3)$p.value, digits=2)`


## (3) Correlations (Table 4)
```{r echo=FALSE}
corlist <- data.frame(income, commit, numrep, ssid, belong, beluncert)
cortable <- data.frame(round(corr.test(as.matrix(corlist))$ci[,1:3], digits=2), round(corr.test(as.matrix(corlist))$ci[,4], digits=4))
colnames(cortable) <- c("lower","r","upper","p")
cortable
```


## (4) Random Assignment Checks

**(a) Unequal variance test, *t*-test, and *M*s and *SD*s for: income**
```{r echo=FALSE}
var.test(income ~ cond)
t.test(income ~ cond, var.equal = TRUE)
describeBy(income, cond)
```
\
**(b) Unequal variance test, *t*-test, and *M*s and *SD*s for: current GPA**
```{r echo=FALSE}
var.test(cgpa ~ cond)
t.test(cgpa ~ cond, var.equal = TRUE)
describeBy(cgpa, cond)
```


## (5) Main Analyses (Table 5)

### Perceived commitment analyses

#### Omnibus test

**Estimates:**
```{r echo=FALSE}
summary(commit.regress <- lm(commit ~ income_c * cond.num))
```

**CIs:**
```{r echo=FALSE}
round(confint(commit.regress, level = 0.95), digits=2)
```

***f*<sup>2</sup> value (effect size) for interaction term:** `r round((summary(commit.regress)$r.squared-summary(lm(commit ~ income_c + cond.num))$r.squared)/(1-summary(commit.regress)$r.squared), digits=3)`

#### Simple slopes tests

**Simple slopes of condition among low- and high-SES students:**
```{r echo=FALSE}
contrasts(cond)
commit.lmres <- lmres(commit ~ income_c * cond.num, mydata)
summary(commit.slopes <- simpleSlope(commit.lmres, pred="cond.num", mod1="income_c"))
```

**CIs for simple slope among low-SES students:** `r round(commit.slopes$simple_slope[1,1] + c(-1,1)*commit.slopes$simple_slope[1,2]*qt(0.975, commit.slopes$Df), digits=2)`  

**CIs for simple slope among high-SES students:** `r round(commit.slopes$simple_slope[2,1] + c(-1,1)*commit.slopes$simple_slope[2,2]*qt(0.975, commit.slopes$Df), digits=2)`\
\


### Numerical representation analyses

#### Omnibus test

**Estimates:**
```{r echo=FALSE}
summary(numrep.regress <- lm(numrep ~ income_c * cond.num))
```

**CIs:**
```{r echo=FALSE}
round(confint(numrep.regress, level = 0.95), digits=2)
```

***f*<sup>2</sup> value (effect size) for interaction term:** `r round((summary(numrep.regress)$r.squared-summary(lm(numrep ~ income_c + cond.num))$r.squared)/(1-summary(numrep.regress)$r.squared), digits=3)`

#### Simple slopes tests

**Simple slopes of condition among low- and high-SES students:**
```{r echo=FALSE}
contrasts(cond)
numrep.lmres <- lmres(numrep ~ income_c * cond.num, mydata)
summary(numrep.slopes <- simpleSlope(numrep.lmres, pred="cond.num", mod1="income_c"))
```

**CIs for simple slope among low-SES students:** `r round(numrep.slopes$simple_slope[1,1] + c(-1,1)*numrep.slopes$simple_slope[1,2]*qt(0.975, numrep.slopes$Df), digits=2)`  

**CIs for simple slope among high-SES students:** `r round(numrep.slopes$simple_slope[2,1] + c(-1,1)*numrep.slopes$simple_slope[2,2]*qt(0.975, numrep.slopes$Df), digits=2)`\
\


### SSID analyses

#### Omnibus test

**Estimates:**
```{r echo=FALSE}
summary(ssid.regress <- lm(ssid ~ income_c * cond.num))
```

**CIs:**
```{r echo=FALSE}
round(confint(ssid.regress, level = 0.95), digits=2)
```

***f*<sup>2</sup> value (effect size) for interaction term:** `r round((summary(ssid.regress)$r.squared-summary(lm(ssid ~ income_c + cond.num))$r.squared)/(1-summary(ssid.regress)$r.squared), digits=3)`

#### Simple slopes tests

**Simple slopes of condition among low- and high-SES students:**
```{r echo=FALSE}
contrasts(cond)
ssid.lmres <- lmres(ssid ~ income_c * cond.num, mydata)
summary(ssid.slopes <- simpleSlope(ssid.lmres, pred="cond.num", mod1="income_c"))
```

**CIs for simple slope among low-SES students:** `r round(ssid.slopes$simple_slope[1,1] + c(-1,1)*ssid.slopes$simple_slope[1,2]*qt(0.975, ssid.slopes$Df), digits=2)`  

**CIs for simple slope among high-SES students:** `r round(ssid.slopes$simple_slope[2,1] + c(-1,1)*ssid.slopes$simple_slope[2,2]*qt(0.975, ssid.slopes$Df), digits=2)`\
\


### Belonging uncertainty analyses

#### Omnibus test

**Estimates:**
```{r echo=FALSE}
summary(beluncert.regress <- lm(beluncert ~ income_c * cond.num))
```

**CIs:**
```{r echo=FALSE}
round(confint(beluncert.regress, level = 0.95), digits=2)
```

***f*<sup>2</sup> value (effect size) for interaction term:** `r round((summary(beluncert.regress)$r.squared-summary(lm(beluncert ~ income_c + cond.num))$r.squared)/(1-summary(beluncert.regress)$r.squared), digits=3)`

#### Simple slopes tests

**Simple slopes of condition among low- and high-SES students:**
```{r echo=FALSE}
contrasts(cond)
beluncert.lmres <- lmres(beluncert ~ income_c * cond.num, mydata)
summary(beluncert.slopes <- simpleSlope(beluncert.lmres, pred="cond.num", mod1="income_c"))
```

**CIs for simple slope among low-SES students:** `r round(beluncert.slopes$simple_slope[1,1] + c(-1,1)*beluncert.slopes$simple_slope[1,2]*qt(0.975, beluncert.slopes$Df), digits=2)`  

**CIs for simple slope among high-SES students:** `r round(beluncert.slopes$simple_slope[2,1] + c(-1,1)*beluncert.slopes$simple_slope[2,2]*qt(0.975, beluncert.slopes$Df), digits=2)`\
\


### Belonging analyses

#### Omnibus test

**Estimates:**
```{r echo=FALSE}
summary(belong.regress <- lm(belong ~ income_c * cond.num))
```

**CIs:**
```{r echo=FALSE}
round(confint(belong.regress, level = 0.95), digits=2)
```

***f*<sup>2</sup> value (effect size) for interaction term:** `r round((summary(belong.regress)$r.squared-summary(lm(belong ~ income_c + cond.num))$r.squared)/(1-summary(belong.regress)$r.squared), digits=3)`

#### Simple slopes tests

**Simple slopes of condition among low- and high-SES students:**
```{r echo=FALSE}
contrasts(cond)
belong.lmres <- lmres(belong ~ income_c * cond.num, mydata)
summary(belong.slopes <- simpleSlope(belong.lmres, pred="cond.num", mod1="income_c"))
```

**CIs for simple slope among low-SES students:** `r round(belong.slopes$simple_slope[1,1] + c(-1,1)*belong.slopes$simple_slope[1,2]*qt(0.975, belong.slopes$Df), digits=2)`  

**CIs for simple slope among high-SES students:** `r round(belong.slopes$simple_slope[2,1] + c(-1,1)*belong.slopes$simple_slope[2,2]*qt(0.975, belong.slopes$Df), digits=2)`\
\


### Analyses with subjective SES (Footnote 2)

### Perceived commitment analyses
```{r echo=FALSE}
summary(lm(commit ~ ladder_c * cond.num))
```

### Numerical representation analyses
```{r echo=FALSE}
summary(lm(numrep ~ ladder_c * cond.num))
```

### SSID analyses
```{r echo=FALSE}
summary(lm(ssid ~ ladder_c * cond.num))
```

### Belonging uncertainty analyses
```{r echo=FALSE}
summary(lm(beluncert ~ ladder_c * cond.num))
```

### Belonging analyses
```{r echo=FALSE}
summary(lm(belong ~ ladder_c * cond.num))
```

```{r include=FALSE, warning=FALSE}
## create and export tables

# correlation (Table 4)
corstarsl <- function(x, docname){ 
  require(Hmisc)
  require(psychometric)
  require(rtf)
  x <- as.matrix(x)
  R <- rcorr(x)$r
  p <- rcorr(x)$P
  mystars <- ifelse(p < .001, "***", ifelse(p < .01, "**", ifelse(p < .05, "*", ifelse(p < .1, "#", ""))))
  R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
  Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
  for (i in 1:length(rcorr(x)$r)) {
    ci <- round(CIr(r = corr.test(as.matrix(corlist))$r[i], n = corr.test(as.matrix(corlist))$n[i], level = .95), 2)
    Rnew[i] <- paste(Rnew[i], " [", ci[1], ", ", ci[2], "]", sep="")
  }
  diag(Rnew) <- paste(diag(R), " ", sep="")
  rownames(Rnew) <- colnames(x)
  colnames(Rnew) <- paste(colnames(x), "", sep="")
  Rnew <- as.matrix(Rnew)
  Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
  Rnew <- as.data.frame(Rnew)
  Rnew <- cbind(Rnew[1:length(Rnew)-1])
  rtffile <- RTF(file=paste(docname, ".doc", sep=""), width=11, height=8.5, font.size=11)
  addTable(rtffile, cbind(rownames(Rnew), Rnew), col.justify="C", header.col.justify="C")
  addText(rtffile, "# p < .10. * p < .05. ** p < .01. *** p < .001.")
  done(rtffile)
}
corstarsl(corlist, "Table 4 (correlations)")

# Table 5 (full model section)
regtable <- function(modellist, docname) {
  sections <- list(NA)
  colnam <- vector("list", length(modellist)*4)
  for (i in 1:length(modellist)) {
    assign(paste("b", i, sep = ""), signif(coef(summary(modellist[[i]])), 3))
    assign(paste("ci", i, sep = ""), paste(get(paste("b", i, sep=""))[, 1], " [", signif(confint(modellist[[i]], level = 0.95)[,1], 3), ", ", signif(confint(modellist[[i]], level = 0.95)[,2], 3), "]", sep = ""))
    assign(paste("t", i, sep = ""), signif(summary(modellist[[i]])$coefficients[,3], 3))
    assign(paste("df", i, sep = ""), rep(summary(modellist[[i]])$df[2],each=4))
    assign(paste("p", i, sep = ""), signif(summary(modellist[[i]])$coefficients[,4], 3))
    assign(paste("section", i, sep = ""), rbind(cbind(get(paste("ci", i, sep = "")), get(paste("t", i, sep = "")), get(paste("df", i, sep = "")), get(paste("p", i, sep = "")))))
    sections[[i]] <- get(paste("section", i, sep = ""))
    colnam[1+(4*(i-1))] <- paste(all.vars(summary(modellist[[i]])$call)[1], ": b (95% CIs)", sep="")
    colnam[2+(4*(i-1))] <- "t"
    colnam[3+(4*(i-1))] <- "df"
    colnam[4+(4*(i-1))] <- "p"
  }
  complete <- do.call(cbind, sections)
  colnames(complete) <- colnam
  rownames(complete) <- c("Intercept", labels(modellist[[i]])[1], labels(modellist[[i]])[2], labels(modellist[[i]])[3])
  complete <- complete[-1,]
  rtffile <- RTF(file=paste(docname, ".doc", sep=""), width=11, height=8.5, font.size=8)  # this can be an .rtf or a .doc
  addTable(rtffile, cbind(rownames(complete), complete), col.justify="C", header.col.justify="C")
  done(rtffile)
}
models <- list(commit.regress, numrep.regress, ssid.regress, beluncert.regress, belong.regress)
regtable(models, "Table 5 (omnibus)")


# Table 5 (simple slopes section)
sstable <- function(modellist, docname) {
  sections <- list(NA)
  colnam <- vector("list", length(modellist)*4)
  for (i in 1:length(modellist)) {
    assign(paste("b", i, sep = ""), signif(summary(modellist[[i]])$simple_slope))
    assign(paste("ci", i, sep = ""), paste(get(paste("b", i, sep=""))[, 1], " [", c(signif(get(paste("b", i, sep=""))[1,1] + c(-1,1)*get(paste("b", i, sep=""))[1,2]*qt(0.975, modellist[[i]]$Df), 3)[1], signif(get(paste("b", i, sep=""))[2,1] + c(-1,1)*get(paste("b", i, sep=""))[2,2]*qt(0.975, modellist[[i]]$Df), 3)[1]), ", ", c(signif(get(paste("b", i, sep=""))[1,1] + c(-1,1)*get(paste("b", i, sep=""))[1,2]*qt(0.975, modellist[[i]]$Df), 3)[2], signif(get(paste("b", i, sep=""))[2,1] + c(-1,1)*get(paste("b", i, sep=""))[2,2]*qt(0.975, modellist[[1]]$Df), 3)[2]), "]", sep = ""))
    assign(paste("t", i, sep = ""), signif(get(paste("b", i, sep=""))[,3], 3))
    assign(paste("df", i, sep = ""), rep(summary(modellist[[i]])$Df,each=2))
    assign(paste("p", i, sep = ""), signif(get(paste("b", i, sep=""))[,4], 3))
    assign(paste("section", i, sep = ""), rbind(cbind(get(paste("ci", i, sep = "")), get(paste("t", i, sep = "")), get(paste("df", i, sep = "")), get(paste("p", i, sep = "")))))
    sections[[i]] <- get(paste("section", i, sep = ""))
    colnam[1+(4*(i-1))] <- paste(modellist[[i]]$nomY, ": b (95% CIs)", sep="")
    colnam[2+(4*(i-1))] <- "t"
    colnam[3+(4*(i-1))] <- "df"
    colnam[4+(4*(i-1))] <- "p"
  }
  complete <- do.call(cbind, sections)
  colnames(complete) <- colnam
  rownames(complete) <- c(row.names(summary(modellist[[i]])$Points)[1], row.names(summary(modellist[[i]])$Points)[2])
  rtffile <- RTF(file=paste(docname, ".doc", sep=""), width=11, height=8.5, font.size=8)
  addTable(rtffile, cbind(rownames(complete), complete), col.justify="C", header.col.justify="C")
  done(rtffile)
}
models.ss <- list(commit.slopes, numrep.slopes, ssid.slopes, beluncert.slopes, belong.slopes)
sstable(models.ss, "Table 5 (simple slopes)")
```
