---
title: "Experiment 1"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: default
---

## (1) Description of Data File

```{r include=FALSE}
options(width=100)

library(RCurl)
library(data.table)
library(pequod)
library(psych)
library(rtf)

## DATA PREPARATION ##

# import data
csv <- getURL("https://raw.githubusercontent.com/abrowman/climate-pspb2016/master/Experiment%201/Climate_Experiment1.csv")
mydata <- read.csv(textConnection(csv))
attach(mydata)
```

The posted data for this study consists of 1 compiled, deidentified data file (total *N* recruited = `r nrow(mydata)`). No data were excluded.

```{r include=FALSE}
# clean data
Income <- gsub("5-6", "5.5", Income)
Income <- as.numeric(gsub("\\.", "", Income))
gGPA <- gsub("3.4-3.5", "3.45", gGPA)
gGPA <- gsub("3.75 \\(hopefully\\!\\)", "3.75", gGPA)
gGPA <- gsub("4.8", "3.8", gGPA)
gGPA <- as.numeric(gGPA)
cGPA <- gsub("nyob", "", cGPA)
cGPA <- gsub("3.7\\?", "3.7", cGPA)
cGPA <- as.numeric(gsub("4.5", "3.5", cGPA))
cGPA[54] <- 3.425

# score measures
cond <- factor(ifelse(!is.na(Item1I), "chilly", ifelse(!is.na(Item1C), "warm" ,NA)), levels=c("warm", "chilly"))
cond.num <- ifelse(cond=="chilly", -1, ifelse(cond=="warm", 1, NA))
options(contrasts=c("contr.sum", "contr.poly"))
eff <- (eff1+eff2+eff3+eff4+eff5)/5
income <- Income
ladder <- US.ldr
ggpa <- gGPA
cgpa <- cGPA

# center continuous predictors
income_c <- scale(income,scale=F,center=T)
ladder_c <- scale(ladder,scale=F,center=T)
```

## (2) Descriptive Statistics (Table 1)

**M~age~ (*SD*):** `r round(mean(Age, na.rm=T), digits=1)` (`r round(sd(Age, na.rm=T), digits=2)`)

**Percentage by gender:**
```{r echo=FALSE}
gender <- factor(ifelse(Gender==1, "male", ifelse(Gender==2, "female", NA)))
round(sort((table(gender, useNA="always")/nrow(mydata)*100),decreasing=F), digits=1)
```

**Percentage by race:**
```{r echo=FALSE}
race <- ifelse(Race==1, "Native-American", ifelse(Race==2, "Asian", ifelse(Race==3, "Black", ifelse(Race==4, "Latino", ifelse(Race==5, "Middle-Eastern", ifelse(Race==6, "Islander", ifelse(Race==7, "White", ifelse(Race==8 | Race=="7,8", "Multi-racial", NA)))))))) # sample by race: 1=native-american; 2=asian; 3=black; 4=latino; 5=middle eastern; 6=islander; 7=white; 8=multi-racial; blank=unreported
round(sort((table(race, useNA="always")/nrow(mydata))*100,decreasing=T), digits=1)
```

**Percentage by class:**
```{r echo=FALSE}
class <- ifelse(Class==1, "Freshman", ifelse(Class==2, "Sophomore", ifelse(Class==3, "Junior", ifelse(Class==4, "Senior", ifelse(Class==5, "Graduate", NA)))))
round(sort((table(class, useNA="always")/nrow(mydata))*100,decreasing=T), digits=1)
```

**Percentage by income:**
```{r echo=FALSE}
income.label <- ifelse(income==1, "(1) $25,000 or less", ifelse(income==2, "(2) $25,001-$40,000", ifelse(income==3, "(3) $40,001-$70,000", ifelse(income==4, "(4) $70,001-$90,000", ifelse(income==5, "(5) $90,001-$120,000", ifelse(income==6, "(6) $120,001-$150,000", ifelse(income==7, "(7) $150,001-$200,000", ifelse(income==8, "(8) $200,001-$300,000", ifelse(income==9, "(9) $300,001 or more", NA)))))))))
round((table(income.label, useNA="always")/nrow(mydata))*100, digits=1)
```

***M*s and *SD*s for income, efficacy (eff), expected graduating GPA (ggpa), and current GPA (cgpa):**  

*Note:* The *M* and *SD* for current GPA are slightly misreported in the published paper due to typoes. Correcting those values (as done here) does not change any conclusions or central values discussed in the paper.
```{r echo=FALSE}
describe(data.frame(income, eff, ggpa, cgpa))[,2:4]
```

**$\alpha$~efficacy~:** `r round(alpha(data.frame(eff1, eff2, eff3, eff4, eff5))$total$raw_alpha, digits=2)`

## (3) Random Assignment Checks

**(a) Unequal variance test, *t*-test, and *M*s and *SD*s for: income**
```{r echo=FALSE}
var.test(income ~ cond)
t.test(income ~ cond, var.equal=TRUE)
describeBy(income, cond)
```
\
**(b) Unequal variance test, *t*-test, and *M*s and *SD*s for: current GPA**
```{r echo=FALSE}
var.test(cgpa ~ cond)
t.test(cgpa ~ cond, var.equal=TRUE)
describeBy(cgpa, cond)
```


## (4) Main Analyses

### Regressing efficacy on condition, income, and their interaction (Table 2)

#### Omnibus test

**Estimates:**
```{r echo=FALSE}
summary(eff.income.regress <- lm(eff ~ income_c * cond.num))
```

**CIs:**
```{r echo=FALSE}
round(confint(eff.income.regress, level = 0.95), digits=2)
```

***f*<sup>2</sup> value (effect size) for interaction term:** `r round((summary(eff.income.regress)$r.squared-summary(lm(eff ~ income_c + cond.num))$r.squared)/(1-summary(eff.income.regress)$r.squared), digits=3)`


#### Simple slopes tests

**Simple slopes of condition among low- and high-SES students:**
```{r echo=FALSE}
contrasts(cond)
eff.income.lmres <- lmres(eff ~ cond.num * income_c, mydata)
summary(eff.income.slopes <- simpleSlope(eff.income.lmres, pred="cond.num", mod1="income_c"))
```

**CIs for simple slope among low-SES students:** `r round(eff.income.slopes$simple_slope[1,1] + c(-1,1)*eff.income.slopes$simple_slope[1,2]*qt(0.975, eff.income.slopes$Df), digits=2)`  

**CIs for simple slope among high-SES students:** `r round(eff.income.slopes$simple_slope[2,1] + c(-1,1)*eff.income.slopes$simple_slope[2,2]*qt(0.975, eff.income.slopes$Df), digits=2)`\
\

**Simple slope of income in each condition (Footnote 3):**
```{r echo=FALSE}
contrasts(cond)
summary(eff.cond.slopes <- simpleSlope(eff.income.lmres, pred="income_c", mod1="cond.num"))
```

**CIs for simple slope among students in the chilly climate condition:** `r round(eff.cond.slopes$simple_slope[1,1] + c(-1,1)*eff.cond.slopes$simple_slope[1,2]*qt(0.975, eff.cond.slopes$Df), digits=2)`  
**CIs for simple slope among students in the warm climate condition:** `r round(eff.cond.slopes$simple_slope[2,1] + c(-1,1)*eff.cond.slopes$simple_slope[2,2]*qt(0.975, eff.cond.slopes$Df), digits=2)`


### Regressing expected graduating GPA on condition, income, and their interaction (Table 2)

#### Omnibus test

**Estimates:**
```{r echo=FALSE}
summary(ggpa.income.regress <- lm(ggpa ~ income_c * cond.num))
```

**CIs:**
```{r echo=FALSE}
round(confint(ggpa.income.regress, level = 0.95), digits=3)
```

***f*<sup>2</sup> value (effect size) for interaction term:** `r round((summary(ggpa.income.regress)$r.squared-summary(lm(ggpa ~ income_c + cond.num))$r.squared)/(1-summary(ggpa.income.regress)$r.squared), digits=3)`

#### Simple slopes tests

**Simple slopes of condition among low- and high-SES students:**
```{r echo=FALSE, warning=FALSE}
contrasts(cond)
ggpa.income.lmres <- lmres(ggpa ~ cond.num * income_c, mydata)
summary(ggpa.income.slopes <- simpleSlope(ggpa.income.lmres, pred="cond.num", mod1="income_c"))
```

**CIs for simple slope among low-SES students:** `r round(ggpa.income.slopes$simple_slope[1,1] + c(-1,1)*ggpa.income.slopes$simple_slope[1,2]*qt(0.975, ggpa.income.slopes$Df), digits=2)`  

**CIs for simple slope among high-SES students:** `r round(ggpa.income.slopes$simple_slope[2,1] + c(-1,1)*ggpa.income.slopes$simple_slope[2,2]*qt(0.975, ggpa.income.slopes$Df), digits=2)`\
\

**Simple slopes of income in each condition (Footnote 3):**
```{r echo=FALSE, warning=FALSE}
contrasts(cond)
summary(ggpa.cond.slopes <- simpleSlope(ggpa.income.lmres, pred="income_c", mod1="cond.num"))
```

**CIs for simple slope among students in the chilly climate condition:** `r round(ggpa.cond.slopes$simple_slope[1,1] + c(-1,1)*ggpa.cond.slopes$simple_slope[1,2]*qt(0.975, ggpa.cond.slopes$Df), digits=3)`  

**CIs for simple slope among students in the warm climate condition:** `r round(ggpa.cond.slopes$simple_slope[2,1] + c(-1,1)*ggpa.cond.slopes$simple_slope[2,2]*qt(0.975, ggpa.cond.slopes$Df), digits=3)`


### Analyses with subjective SES (Footnote 2)

**For efficacy:**
```{r echo=FALSE}
summary(eff.ladder.regress <- lm(eff ~ ladder_c * cond.num))
```

**For expected graduating GPA:**
```{r echo=FALSE}
summary(ggpa.ladder.regress <- lm(ggpa ~ ladder_c * cond.num))
```


## (5) Figures

### Regressing efficacy on condition, income, and their interaction (Figure 1a)
```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=6}
eff.plot.points <- data.frame(c("Chilly climate condition","Warm climate condition","Chilly climate condition","Warm climate condition"), c("Low-SES","Low-SES","High-SES","High-SES"), c(eff.income.slopes$Points[1,1], eff.income.slopes$Points[1,2], eff.income.slopes$Points[2,1], eff.income.slopes$Points[2,2]))
colnames(eff.plot.points) <- c("cond","SES","score")

eff.plot <- ggplot(data=eff.plot.points, aes(x=cond, y=score, linetype=SES)) +
  geom_line(aes(group = SES)) +
  theme_bw(base_family = "Arial", base_size = 16) +
  theme(legend.position = c(.18,.94)) +
  theme(legend.key = element_blank()) +
  theme(legend.title = element_blank()) +
  scale_linetype_discrete(breaks=c("Low-SES", "High-SES"), labels=c("Low-SES students (-1 SD)", "High-SES students (+1 SD)")) +
  theme(axis.title.x = element_blank()) +
  theme(axis.title.y = element_text(vjust=1.3)) +
  theme(panel.grid.minor = element_blank()) +
  theme(panel.grid.major = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(axis.line = element_line(color = 'black')) +
  coord_cartesian(ylim=c(4, 6)) +
  scale_y_continuous("Academic efficacy", breaks = seq(4, 6, by = .5)) +
  annotate("text", x=2.5, y=5.97, label="(a)")
eff.plot

ggsave('Figure 1a.png', width=9, height=6, unit='in', dpi=300)
```

### Regressing expected graduating GPA on condition, income, and their interaction (Figure 1b)
```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=6}
ggpa.plot.points <- data.frame(c("Chilly climate condition","Warm climate condition","Chilly climate condition","Warm climate condition"), c("Low-SES","Low-SES","High-SES","High-SES"), c(ggpa.income.slopes$Points[1,1], ggpa.income.slopes$Points[1,2], ggpa.income.slopes$Points[2,1], ggpa.income.slopes$Points[2,2]))
colnames(ggpa.plot.points) <- c("cond","SES","score")

ggpa.plot <- ggplot(data=ggpa.plot.points, aes(x=cond, y=score, linetype=SES)) + # data to plot
  geom_line(aes(group = SES)) +
  theme_bw(base_family = "Arial", base_size = 16) +
  theme(legend.position = "none") +
  scale_linetype_discrete(breaks=c("Low-SES", "High-SES"), labels=c("Low-SES students (-1 SD)", "High-SES students (+1 SD)")) +
  theme(axis.title.x = element_blank()) +
  theme(axis.title.y = element_text(vjust=1.3)) +
  theme(panel.grid.minor = element_blank()) +
  theme(panel.grid.major = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(axis.line = element_line(color = 'black')) +
  coord_cartesian(ylim=c(3.2, 3.8)) +
  scale_y_continuous("Expected graduating GPA", breaks = seq(3.2, 3.8, by = .1)) +
  annotate("text", x=2.5, y=3.79, label="(b)")
ggpa.plot

ggsave('Figure 1b.png', width=9, height=6, unit='in', dpi=300)
```

```{r include=FALSE, warning=FALSE}
## Scripts for auto-producing and exporting regression tables to Word format

# Omnibus regression results (Table 2)
regtable <- function(modellist, docname) {
  sections <- list(NA)
  colnam <- vector("list", length(modellist)*4)
  for (i in 1:length(modellist)) {
    assign(paste("b", i, sep = ""), signif(coef(summary(modellist[[i]])), 3))
    assign(paste("ci", i, sep = ""), paste(get(paste("b", i, sep=""))[, 1], " [", signif(confint(modellist[[i]], level = 0.95)[,1], 3), ", ", signif(confint(modellist[[i]], level = 0.95)[,2], 3), "]", sep = ""))
    assign(paste("t", i, sep = ""), signif(summary(modellist[[i]])$coefficients[,3], 3))
    assign(paste("df", i, sep = ""), rep(summary(modellist[[i]])$df[2],each=4))
    assign(paste("p", i, sep = ""), signif(summary(modellist[[i]])$coefficients[,4], 3))
    assign(paste("section", i, sep = ""), rbind(cbind(get(paste("ci", i, sep = "")), get(paste("t", i, sep = "")), get(paste("df", i, sep = "")), get(paste("p", i, sep = "")))))
    sections[[i]] <- get(paste("section", i, sep = ""))
    colnam[1+(4*(i-1))] <- paste(all.vars(summary(modellist[[i]])$call)[1], ": b (95% CIs)", sep="")
    colnam[2+(4*(i-1))] <- "t"
    colnam[3+(4*(i-1))] <- "df"
    colnam[4+(4*(i-1))] <- "p"
  }
  complete <- do.call(cbind, sections)
  colnames(complete) <- colnam
  rownames(complete) <- c("Intercept", labels(modellist[[i]])[1], labels(modellist[[i]])[2], labels(modellist[[i]])[3])
  complete <- complete[-1,]
  rtffile <- RTF(file=paste(docname, ".doc", sep=""), width=11, height=8.5, font.size=12)
  addTable(rtffile, cbind(rownames(complete), complete), col.justify="C", header.col.justify="C")
  done(rtffile)
}
models <- list(eff.income.regress, ggpa.income.regress)
regtable(models, "Table 2a")

# Simple slope analysis results (Table 2)
sstable <- function(modellist, docname) {
  sections <- list(NA)
  colnam <- vector("list", length(modellist)*4)
  for (i in 1:length(modellist)) {
    assign(paste("b", i, sep = ""), signif(summary(modellist[[i]])$simple_slope))
    assign(paste("ci", i, sep = ""), paste(get(paste("b", i, sep=""))[, 1], " [", c(signif(get(paste("b", i, sep=""))[1,1] + c(-1,1)*get(paste("b", i, sep=""))[1,2]*qt(0.975, modellist[[i]]$Df), 3)[1], signif(get(paste("b", i, sep=""))[2,1] + c(-1,1)*get(paste("b", i, sep=""))[2,2]*qt(0.975, modellist[[i]]$Df), 3)[1]), ", ", c(signif(get(paste("b", i, sep=""))[1,1] + c(-1,1)*get(paste("b", i, sep=""))[1,2]*qt(0.975, modellist[[i]]$Df), 3)[2], signif(get(paste("b", i, sep=""))[2,1] + c(-1,1)*get(paste("b", i, sep=""))[2,2]*qt(0.975, modellist[[1]]$Df), 3)[2]), "]", sep = ""))
    assign(paste("t", i, sep = ""), signif(get(paste("b", i, sep=""))[,3], 3))
    assign(paste("df", i, sep = ""), rep(summary(modellist[[i]])$Df,each=2))
    assign(paste("p", i, sep = ""), signif(get(paste("b", i, sep=""))[,4], 3))
    assign(paste("section", i, sep = ""), rbind(cbind(get(paste("ci", i, sep = "")), get(paste("t", i, sep = "")), get(paste("df", i, sep = "")), get(paste("p", i, sep = "")))))
    sections[[i]] <- get(paste("section", i, sep = ""))
    colnam[1+(4*(i-1))] <- paste(modellist[[i]]$nomY, ": b (95% CIs)", sep="")
    colnam[2+(4*(i-1))] <- "t"
    colnam[3+(4*(i-1))] <- "df"
    colnam[4+(4*(i-1))] <- "p"
  }
  complete <- do.call(cbind, sections)
  colnames(complete) <- colnam
  rownames(complete) <- c(row.names(summary(modellist[[i]])$Points)[1], row.names(summary(modellist[[i]])$Points)[2])
  rtffile <- RTF(file=paste(docname, ".doc", sep=""), width=11, height=8.5, font.size=12)
  addTable(rtffile, cbind(rownames(complete), complete), col.justify="C", header.col.justify="C")
  done(rtffile)
}
models.ss <- list(eff.income.slopes, ggpa.income.slopes)
sstable(models.ss, "Table 2b")
```
